---
title: "Assignment 5"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
p_load(tidyverse, brms)
```


# Data Preparation
```{r Loading and Scaling Data}
# load the data
d <- read_csv("Ass4_data.csv", col_types = cols(ID = col_character()))

d$npause[d$npause == 0] <- NA
d$`speechrate (nsyll/dur)`[d$`speechrate (nsyll/dur)` == 0.0] <- NA
d$pauseFreq <- d$Duration / d$npause
d$speechrate <-d$`speechrate (nsyll/dur)`

# Selecting data
d <- d %>% 
  subset(Language == "dk") %>% 
  select(ID, Diagnosis, Pitch_IQR, Pitch_Median, Pitch_MAD, Pitch_SD, Pitch_Mean, pauseFreq, speechrate)
  
# Making diagnosis a factor with levels 0 and 1
d$Diagnosis <- as.factor(d$Diagnosis)
levels(d$Diagnosis) <- c('0', '1')



# Making a version where the data is scaled between 0-1 after being grouped
# Grouping scaled data
grouped_scaled_01 <- d %>% 
  group_by(ID) %>% 
  summarise(Pitch_IQR = mean(Pitch_IQR),
            Pitch_Median = mean(Pitch_Median),
            Pitch_Mean = mean(Pitch_Mean),
            Pitch_MAD = mean(Pitch_MAD),
            Pitch_SD = mean(Pitch_SD),
            pauseFreq = mean(pauseFreq, na.rm = T),
            speechrate = mean(speechrate, na.rm = T),
            Diagnosis = Diagnosis[1])

grouped_scaled_01 <- grouped_scaled_01 %>% 
  mutate(Pitch_IQR = scale0_1(log(Pitch_IQR)),
         Pitch_MAD = scale0_1(log(Pitch_MAD)),
         Pitch_Median = scale0_1(Pitch_Median),
         Pitch_Mean = scale0_1(Pitch_Mean),
         Pitch_SD = scale0_1(Pitch_SD),
         pauseFreq = scale0_1(log(pauseFreq), na.rm = T),
         speechrate = scale0_1(speechrate, na.rm = T))%>%
  select(ID, Diagnosis, Pitch_IQR, Pitch_MAD, Pitch_Median, Pitch_SD, Pitch_Mean, pauseFreq, speechrate)

# Checking dist of data grouped by participants who have been scaled between 0-1
histogram_plot(grouped_scaled_01, grouped_scaled_01$pauseFreq, 0.05, "Frequency", "Pause frequency")
histogram_plot(grouped_scaled_01, grouped_scaled_01$Pitch_MAD, 0.05, "Frequency", "MAD")

# Write csv with new dataset
write_csv(grouped_scaled_01, "grouped_scaled_01.csv")
```



# Data Preparation
```{r Maniupulating Data}
# Loading in the original data
original_data <- read_csv("grouped_scaled_01.csv")

original_data$diagnosis <- as.factor(original_data$Diagnosis)

original_bv <- original_data %>% select(ID, diagnosis, Pitch_IQR, Pitch_MAD, pauseFreq, speechrate)


# Loading in simulated data and grouping it together
sim_data <- read_csv("batch_1000_1000_new.csv")

grouped_sim_data <- sim_data %>% 
  group_by(AgentId) %>% 
  summarise(activity = mean(activity), change_IQR = mean(change_IQR),
            change_MAD = mean(change_MAD), change_PauseFreq = mean(change_PauseFreq),
            change_Speechrate = mean(change_Speechrate), conversation_time = mean(conversation_time),
            interaction_time = mean(interaction_time), interactions = mean(interactions),
            mad = mean(mad), pauseFreq = mean(pauseFreq), speechrate = mean(speechrate), 
            abs_IQR = mean(abs_change_IQR), abs_MAD = mean(abs_change_MAD), IQR = mean(IQR),
            speechrate = mean(speechrate), mad = mean(mad), pauseFreq = mean(pauseFreq),
            diagnosis = diagnosis[1])


# Add ID column
colnames(grouped_sim_data)[1] <- "ID"


# set diagnosis as factor with levels 0 and 1
grouped_sim_data$diagnosis <- as.factor(grouped_sim_data$diagnosis)
levels(grouped_sim_data$diagnosis) <- c('0', '1')


# Scaling simulation data between 0 - 1
grouped_sim_data <- grouped_sim_data %>% 
  mutate(activity = scale(activity), change_IQR = scale(change_IQR), change_MAD = scale(change_MAD),
         change_PauseFreq = scale(change_PauseFreq), change_Speechrate = scale(change_Speechrate),
         conversation_time = scale(conversation_time), interaction_time = scale(interaction_time),
         interactions = scale(interactions), abs_IQR = scale(abs_IQR), abs_MAD = scale(abs_MAD))

```


# Specifying Bayesian Formulas for Original Data
```{r}
IQR_f <- bf(Pitch_IQR ~ 1 + diagnosis)
MAD_f <- bf(Pitch_MAD ~ 1 + diagnosis)
pauseFreq_f <- bf(pauseFreq ~ 1 + diagnosis)
speechrate_f <- bf(speechrate ~ 1 + diagnosis)

get_prior(IQR_f, data = original_data)

original_prior <- c(
  prior(normal(0, .2), class = b),
  prior(normal(.5, .1), class = Intercept),
  prior(normal(.1, .5), class = sigma)
)
```


Modeling Original Variables
```{r}
# INTER-QUARTILE RANGE - AKA PITCH VARIABILITY
orig_IQR_pc <- brm(
  IQR_f,
  data = original_data,
  family = gaussian(),
  prior = original_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(orig_IQR_pc, nsamples = 100)

orig_IQR_m <- brm(
  IQR_f,
  data = original_data,
  family = gaussian(),
  prior = original_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(orig_IQR_m, nsamples = 100)

summary(orig_IQR_m)
hypothesis(orig_IQR_m, "diagnosis1 < 0")


# MEAN ABSOLUTE STANDARD DEVIATION
orig_MAD_pc <- brm(
  MAD_f,
  data = original_data,
  family = gaussian(),
  prior = original_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(orig_MAD_pc, nsamples = 100)

orig_MAD_m <- brm(
  MAD_f,
  data = original_data,
  family = gaussian(),
  prior = original_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(orig_MAD_m, nsamples = 100)

summary(orig_MAD_m)
hypothesis(orig_MAD_m, "diagnosis1 < 0")


# PAUSE FREQUENCY
orig_pause_pc <- brm(
  pauseFreq_f,
  data = original_data,
  family = gaussian(),
  prior = original_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(orig_pause_pc, nsamples = 100)

orig_pause_m <- brm(
  pauseFreq_f,
  data = original_data,
  family = gaussian(),
  prior = original_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(orig_pause_m, nsamples = 100)

summary(orig_pause_m)
hypothesis(orig_pause_m, "diagnosis1 < 0")


# SPEECHRATE
orig_speechrate_pc <- brm(
  speechrate_f,
  data = original_data,
  family = gaussian(),
  prior = original_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(orig_speechrate_pc, nsamples = 100)

orig_speechrate_m <- brm(
  speechrate_f,
  data = original_data,
  family = gaussian(),
  prior = original_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(orig_speechrate_m, nsamples = 100)

summary(orig_speechrate_m)
hypothesis(orig_speechrate_m, "diagnosis1 > 0")
```

The model yields a credible relation between pitch IQR and diagnosis, with TD's showing lower iqr (b = -0.1, CIs = -0.18, -0.02, Evid. Ratio = 34.09).
The model yields a credible relation between mean absolute standard deviation and diagnosis, with TD's showing a lower MAD (b = -0.13, CIs = -0.21, -0.05, Evid. Ratio = 199).
The model yields a credible relation between pause frequency and diagnosis with TD's showing a lower overall pause frequency (b = -0.09, CIs = -0.18, -0.01, Evid. Ratio = 25.32).
The model yields a credible relation between speechrate and diagnosis with TD's showing a higher speechrate (b = 0.09, CIs = 0.01, 0.18, Evid. Ratio = 23.69).



# Specifing Bayesian Formulas for Simulated Data
```{r Simple Bayesian analyses}
# Defining model formulas
changeIQR_f <- bf(change_IQR ~ 1 + diagnosis)
changeMAD_f <- bf(change_MAD ~ 1 + diagnosis)
changePauseFreq_f <- bf(change_PauseFreq ~ 1 + diagnosis)
changeSpeechrate_f <- bf(change_Speechrate ~ 1 + diagnosis)
absIQR_f <- bf(abs_IQR ~ 1 + diagnosis)
absMAD_f <- bf(abs_MAD ~ 1 + diagnosis)
activity_f <- bf(activity ~ 1 + diagnosis)
interactions_f <- bf(interactions ~ 1 + diagnosis)
interactiontime_f <- bf(interaction_time ~ 1 + diagnosis)
conversationtime_f <- bf(conversation_time ~ 1 + diagnosis)


# Getting necessary priors (identical for all models)
get_prior(changeIQR_f, data = grouped_sim_data)


# Setting skeptical priors
sim_prior <- c(
  prior(normal(0, 1), class = b),
  prior(normal(.5, .1), class = Intercept),
  prior(normal(.1, .5), class = sigma)
)
```


Happy Modeling of Simulated Variables!
```{r}
## Change Pitch Variability
# Prior predictive check
IQR_pc <- brm(
  changeIQR_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)

pp_check(IQR_pc, nsamples = 100)

# Fitting the model
IQR_m <- brm(
  changeIQR_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)

pp_check(IQR_m, nsamples = 100)
summary(IQR_m)
hypothesis(IQR_m, "diagnosis1 > 0")



## Change Mean Absolute Standard Deviation
MAD_pc <- brm(
  changeMAD_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(MAD_pc, nsamples = 100)

MAD_m <- brm(
  changeMAD_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(IQR_m, nsamples = 100)

summary(MAD_m)
hypothesis(MAD_m, "diagnosis1 > 0")



## Change Pause Frequency
PauseFreq_pc <- brm(
  changePauseFreq_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(PauseFreq_pc, nsamples = 100)

PauseFreq_m <- brm(
  changePauseFreq_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(PauseFreq_m, nsamples = 100)

summary(PauseFreq_m)
hypothesis(PauseFreq_m, "diagnosis1 > 0")



## Change Speech Rate
speechrate_pc <- brm(
  changeSpeechrate_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(speechrate_pc, nsamples = 100)

speechrate_m <- brm(
  changeSpeechrate_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(speechrate_m, nsamples = 100)

summary(speechrate_m)
hypothesis(speechrate_m, "diagnosis1 > 0")



## Absolute IQR Score
absIQR_pc <- brm(
  absIQR_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(absIQR_pc, nsamples = 100)

absIQR_m <- brm(
  absIQR_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(absIQR_m, nsamples = 100)

summary(absIQR_m)
hypothesis(absIQR_m, "diagnosis1 > 0")



## Absolute Mean Absolute Deviation Score
absMAD_pc <- brm(
  absMAD_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(absMAD_pc, nsamples = 100)

absMAD_m <- brm(
  absMAD_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(absMAD_m, nsamples = 100)

summary(absMAD_m)
hypothesis(absMAD_m, "diagnosis1 > 0")



## Activitiy Score
activity_pc <- brm(
  activity_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(activity_pc, nsamples = 100)

activity_m <- brm(
  activity_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(activity_m, nsamples = 100)

summary(activity_m)
hypothesis(activity_m, "diagnosis1 > 0")



## Overall Interactions
interactions_pc <- brm(
  interactions_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(interactions_pc, nsamples = 100)

interactions_m <- brm(
  interactions_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(interactions_m, nsamples = 100)

summary(interactions_m)
hypothesis(interactions_m, "diagnosis1 > 0")



## Overall Interaction Time
interactiontime_pc <- brm(
  interactiontime_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(interactiontime_pc, nsamples = 100)

interactiontime_m <- brm(
  interactiontime_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(interactiontime_m, nsamples = 100)

summary(interactiontime_m)
hypothesis(interactiontime_m, "diagnosis1 > 0")



## Overall Conversation Time
conversationtime_pc <- brm(
  conversationtime_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(conversationtime_pc, nsamples = 100)

conversationtime_m <- brm(
  conversationtime_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = sim_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(conversationtime_m, nsamples = 100)

summary(conversationtime_m)
hypothesis(conversationtime_m, "diagnosis1 > 0")
```


Hypothesis Testing
```{r Hypotheses}
hypothesis(IQR_m, "diagnosis1 > 0")
hypothesis(MAD_m, "diagnosis1 > 0")
summary(MAD_m)
hypothesis(PauseFreq_m, "diagnosis1 > 0")
hypothesis(speechrate_m, "diagnosis1 < 0")
hypothesis(absIQR_m, "diagnosis1 > 0")
plot(hypothesis(absMAD_m, "diagnosis1 > 0"))
summary(absMAD_m)
hypothesis(activity_m, "diagnosis1 > 0")
hypothesis(interactions_m, "diagnosis1 > 0")
hypothesis(interactiontime_m, "diagnosis1 > 0")
hypothesis(conversationtime_m, "diagnosis1 > 0")
```

Diagnosis produced a credible effect on change in MAD score (b = 0.67, CIs = 0.29, 1.04, Evid. Ratio = Inf).
Diagnosis produced a credible effect on absolute IQR score (b = 0.69, CIs = 0.33, 1.05, Evid. Ratio = 1999).
Diagnosis produced a credible effect on amount of interactions (b = 0.69, CIs = 0.32, 1.05, Evid. Ratio = 666).
Diagnosis produced a credible effect on overall interaction time (b = 0.71, CIs = 0.34, 1.06, Evid. Ratio = 399).


Modeling Simulated Interaction Effects
```{r Interactions}
interaction1_f <- bf(abs_MAD ~ 0 + diagnosis:activity)
interaction2_f <- bf(change_MAD ~ 0 + diagnosis:activity)
interaction3_f <- bf(interactions ~ 0 + diagnosis:abs_MAD)
interaction4_f <- bf(abs_MAD ~ 0 + diagnosis:interactions)

get_prior(interaction1_f, data = grouped_sim_data)

interaction_prior <- c(
  prior(normal(0, 1), class = "b"),
  prior(normal(.5, 1), class = "sigma")
)


interaction1_pc <- brm(
  interaction1_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = interaction_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(interaction1_pc, nsamples = 100)

interaction1_m <- brm(
  interaction1_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = interaction_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(interaction1_m, nsamples = 100)

summary(interaction1_m)
hypothesis(interaction1_m, "diagnosis1:activity > diagnosis0:activity")




interaction2_pc <- brm(
  interaction2_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = interaction_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(interaction2_pc, nsamples = 100)

interaction2_m <- brm(
  interaction2_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = interaction_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(interaction2_m, nsamples = 100)

summary(interaction2_m)
hypothesis(interaction2_m, "diagnosis1:activity < diagnosis0:activity")


interaction3_pc <- brm(
  interaction3_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = interaction_prior,
  sample_prior = "only",
  chains = 2,
  iter = 2000
)
pp_check(interaction3_pc, nsamples = 100)

interaction3_m <- brm(
  interaction3_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = interaction_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(interaction3_m, nsamples = 100)

summary(interaction3_m)
hypothesis(interaction3_m, "diagnosis1:abs_MAD > diagnosis0:abs_MAD")



interaction4_m <- brm(
  interaction4_f,
  data = grouped_sim_data,
  family = gaussian(),
  prior = interaction_prior,
  sample_prior = T,
  chains = 2,
  iter = 2000
)
pp_check(interaction4_m, nsamples = 100)

summary(interaction4_m)
summary(interaction4_m, "diagnosis1:interactions > diagnosis0:interactions")
```

